Silenced in the Chamber: Analyzing Gendered Speech Interruptions in the Australian Parliament (1998–2022)

### 2025-02-28 paper/paper.qmd
more stm

@@

```{r}
#| include: false
#| warning: false
#| message: false
hansard_corpus <- read_parquet(here::here("data/02-analysis_data/cleaned_data.parquet"))
```

```{r}
#| include: false
#| warning: false
#| message: false
# View the first few rows
head(hansard_corpus, 150)

@@

@tbl-ingov1, @tbl-ingov2, and @fig-gov offer a detailed look at interruptions during parliamentary sessions, categorized by members' government status and gender. Notably, interruptions are extremely rare for members with clearly identified status. This indicates a potential correlation between established government status and more orderly interactions, highlighting the importance of role clarity in maintaining decorum.

@@

@@

@@

# Print the data to see the word counts
head(hansard_corpus1)
```

@@

# Create the density plot with boundary outlines
ggplot(filtered_data, aes(x = word_count, fill = gender, color = gender)) +
  geom_density(alpha = 0.5) +
  scale_fill_manual(
    values = c("Female" = "pink", "Male" = "lightblue")
  ) +
  scale_color_manual(
    values = c("Female" = "red", "Male" = "blue")
  ) +
  labs(
    title = "Density Distribution of Word Count by Gender (Word Count ≤ 500)",
    x = "Word Count",
    y = "Density",
    fill = "Gender",
    color = "Gender"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5),  # Center the title
    legend.position = "right"               # Place the legend to the right
  )

@@

model1 <-
  readRDS(file = here::here("models/model1.rds"))

@@

```{r}
#| include: false
#| eval: true
#| warning: false
#| message: false
model2 <-
  readRDS(file = here::here("models/model2.rds"))

@@

>>>>

library(stm)
library(quanteda)
library(tm)
```

```{r}
#| include: false
#| warning: false
#| message: false
hansard_corpus <- read_parquet(here::here("C:/Users/35911/Downloads/Hansard_Interruptions/data/02-analysis_data/cleaned_data.parquet"))

@@

```{r}
#| warning: false
#| message: false
# View the first few rows
#head(hansard_corpus, 150)

@@

@tbl-ingov1 and @fig-gov offer a detailed look at interruptions during parliamentary sessions, categorized by members' government status and gender. Notably, interruptions are extremely rare for members with clearly identified status. This indicates a potential correlation between established government status and more orderly interactions, highlighting the importance of role clarity in maintaining decorum.

@@

# Print the resulting dataset to verify
#print(interruptions_question_analysis)

@@

# Print the resulting dataset to verify
#print(interruptions_question_analysis)
```

@@

# Print the data to see the word counts
#head(hansard_corpus1)
```

@@

# Create the density plot with boundary outlines
ggplot(filtered_data, aes(x = word_count, fill = gender, color = gender)) +
  geom_density(alpha = 0.5) +
  scale_fill_manual(
    values = c("Female" = "pink", "Male" = "lightblue")
  ) +
  scale_color_manual(
    values = c("Female" = "red", "Male" = "blue")
  ) +
  labs(
    title = "Density Distribution of Word Count by Gender (Word Count < 500)",
    x = "Word Count",
    y = "Density",
    fill = "Gender",
    color = "Gender"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5),  # Center the title
    legend.position = "right"               # Place the legend to the right
  )

@@

model1 <-
  readRDS(file = here::here("C:/Users/35911/Downloads/Hansard_Interruptions/models/model1.rds"))

@@

```{r}
#| include: false
#| eval: true
#| warning: false
#| message: false
model2 <-
  readRDS(file = here::here("C:/Users/35911/Downloads/Hansard_Interruptions/models/model2.rds"))

@@

```{r}
#| include: false
#| eval: true
#| warning: false
#| message: false
hansard_topics <- readRDS(file = here::here("C:/Users/35911/Downloads/Hansard_Interruptions/models/hansard_topics.rda"))
```

```{r}
#| include: false
#| warning: false
labelTopics(hansard_topics)
```

To analyze discussions in the Australian Parliament, I applied Structural Topic Modeling (STM) to identify key themes in speeches from 2020. The process began with data preprocessing, where I filtered the dataset to include only relevant entries, removed missing values, and structured the text data for analysis. A document-feature matrix (DFM) was then created, representing word frequencies while eliminating punctuation, stopwords, and low-frequency terms to improve model efficiency.

The STM model was trained with 10 topics and incorporated party affiliation and speaker gender as prevalence covariates, allowing for an examination of how these factors influence topic distribution. The model was optimized over 50 iterations, ensuring convergence and stability. The final output provided a breakdown of the dominant topics, with key terms extracted using highest probability words, FREX words (distinctive to each topic), Lift words (rare but highly relevant), and Score words (best summarizing a topic’s meaning).

```{r, fig.width=10, fig.height=6}
#| echo: false
#| eval: true
#| label: tbl-stm_prop
#| tbl-cap: "STM Topic Proportion Plot "
#| warning: false
plot(hansard_topics, type = "summary", n = 5)
```

@tbl-stm_prop provides an overview of the most prevalent themes in the 2020 Australian parliamentary speeches. The x-axis represents the expected topic proportions, indicating how frequently each topic appears across the dataset. The y-axis lists the top topics, summarized by their most representative words.

From this visualization, Topic 5 ("minister, government, time, just") emerges as the most dominant theme, suggesting that government policies, ministerial discussions, and legislative matters were central to parliamentary discourse. Topic 8 ("people, government, need, now") follows closely, highlighting debates around public needs, government interventions, and social policies. Other significant topics include healthcare (Topic 10), economic measures (Topic 4), legal and family-related issues (Topic 7), and national security (Topic 1).

The distribution of topics provides insights into policy priorities and legislative focus areas during 2020. By incorporating metadata such as speaker affiliation and gender, further analysis can uncover how different political groups or demographics engaged with these key issues.

```{r}
#| include: false
library(stm)
library(dplyr)
library(readr)

# Define years and create empty list for storage
years <- 1998:2022
topic_props_list <- list()

# Loop through each year and extract topic proportions & keywords
for (year in years) {
  model_path <- paste0("C:/Users/35911/Downloads/Hansard_Interruptions/models/hansard_topics_", year, ".rda")

  if (file.exists(model_path)) {
    stm_model <- readRDS(model_path)

    # Extract topic proportions
    topic_props <- data.frame(
      Year = year,
      Topic = paste0("Topic_", 1:ncol(stm_model$theta)),  # STM topics
      Proportion = colMeans(stm_model$theta)  # Average topic proportion
    )

    # Extract top words for each topic
    topic_labels <- labelTopics(stm_model, n = 10)$prob  # Adjust 'n' for more words
    topic_keywords <- apply(topic_labels, 1, paste, collapse = ", ")  # Convert to string

    # Add keywords to dataframe
    topic_props$Keywords <- topic_keywords

    # Store results
    topic_props_list[[as.character(year)]] <- topic_props
  } else {
    print(paste("Skipping", year, "- No Model Found"))
  }
}

# Combine results into one dataframe
topic_summary <- bind_rows(topic_props_list)

# Save extracted topics & keywords
write_csv(topic_summary, "topic_summary_1998_2022.csv")

# View first few rows
#print(topic_summary)

```
```{r}
#| include: false
library(dplyr)
library(tidyr)
library(readr)

# Select the most prominent topic per year
top_topics_per_year <- topic_summary %>%
  group_by(Year) %>%
  summarise(Top_Topic = first(Topic), Keywords = first(Keywords))  # First topic listed is the most prominent

# Display table
#print(top_topics_per_year)

```{r}
#| echo: false
#| eval: true
#| label: tbl-prom_topic
#| tbl-cap: "Most Prominent Topics Per Year"
#| warning: false

library(gt)

# Create a nicely formatted table without the "Top Topic" column
top_topics_per_year %>%
  select(-Top_Topic) %>%  # Remove "Top Topic" column
  gt() %>%
  tab_header(
    title = "Most Prominent Topics Per Year",
    subtitle = "Extracted from Parliamentary Speeches (1998-2022)"
  ) %>%
  cols_label(
    Year = "Year",
    Keywords = "Representative Keywords"
  ) %>%
  fmt_markdown(columns = Keywords) %>%
  tab_options(
    table.font.size = "small",
    heading.title.font.size = "large",
    heading.subtitle.font.size = "medium"
  )

To analyze the evolution of parliamentary discourse over time, I ran the Structural Topic Model (STM) on each year from 1998 to 2022, extracting the most prominent topics discussed in Australian parliamentary speeches annually. This approach allows us to track changes in political priorities and public policy discussions across different governments and time periods.

@tbl-prom_topic that extracted topics from Australian parliamentary speeches from 1998 to 2022 reveal recurring themes in economic policy, governance, and public welfare. In 1998, keywords such as "funding, economy, billion, jobs, growth" indicate a strong focus on financial matters, possibly reflecting the economic policies of the Howard government. By 1999, discussions shifted toward legislative processes, with terms like "committee, country, parliament, act, bill, interest", suggesting an emphasis on governance and parliamentary procedures.

In the early 2000s, taxation and economic regulation became central themes. The year 2000 prominently features "tax, GST, regional, cost", likely indicating debates around the introduction of the Goods and Services Tax (GST), a major economic reform at the time. By 2001, discussions expanded to include "question, education, state, leader, speaker", showing parliamentary focus on education policy and leadership discussions. The presence of "security, report, speaker" in later years, such as 2007, suggests increasing discussions about national security and legislative oversight.

The financial crisis of 2008 appears to have influenced parliamentary discussions, as evidenced by the prominence of "government, economic, parliament, working" in that year. This aligns with government efforts to stabilize the economy and implement stimulus measures. By 2010, speeches featured "policy, education, billion, services", indicating continued investment in education and public services. In 2015, "education, students, infrastructure, training, funding" suggests an ongoing emphasis on higher education and workforce development.

More recently, parliamentary discussions reflected the impact of the COVID-19 pandemic. In 2020, "government, security, legislation, parliament" emerged as dominant themes, which likely relate to emergency public health measures and pandemic response policies. By 2021 and 2022, terms such as "community, health, local, member" indicate a growing focus on public health, social support, and government intervention in economic recovery.

Overall, this analysis provides a clear trajectory of parliamentary priorities over time. While early discussions centered on economic policy and legislative procedures, later years saw a growing emphasis on education, national security, and, more recently, public health. The keywords extracted from parliamentary speeches highlight shifting political and policy concerns, aligning with major national and global events during this period.
```{r}
#| include: false
# Load necessary libraries
library(dplyr)
library(ggplot2)
library(readr)

topic_spikes <- topic_summary %>%
  arrange(Topic, Year) %>%
  group_by(Topic) %>%
  mutate(Change = Proportion - lag(Proportion)) %>%
  filter(!is.na(Change))  # Remove first year (no previous comparison)

# Ensure unique topic-year-keyword mapping
topic_spikes <- topic_spikes %>%
  left_join(topic_summary %>% select(Year, Topic, Keywords), by = c("Year", "Topic")) %>%
  distinct() %>%
  select(-Keywords.x) %>%
  rename(Keywords = Keywords.y)

# Verify that Keywords exist
print(head(topic_spikes))

```{r}
#| echo: false
#| eval: true
#| label: fig-keytopics
#| fig-cap: "Mentions of Key Topics Over Time"
#| warning: false
library(stringr)
library(RColorBrewer)  # Better color schemes

# Define keywords of interest
keywords_of_interest <- c("education", "health", "economy", "policy", "tax", "energy")

# Extract individual keyword counts separately
tracked_topics <- topic_summary %>%
  mutate(Cleaned_Keywords = str_extract_all(Keywords, paste(keywords_of_interest, collapse = "|"))) %>%
  unnest(Cleaned_Keywords) %>%  # This separates each keyword into its own row
  filter(Cleaned_Keywords != "") %>%
  group_by(Year, Cleaned_Keywords) %>%
  summarise(Count = sum(Proportion * 1000), .groups = "drop")  # Count each separately

# Define color palette (use distinct colors)
color_palette <- brewer.pal(n = length(unique(tracked_topics$Cleaned_Keywords)), name = "Dark2")

# Plot evolution over time with improved colors
ggplot(tracked_topics, aes(x = Year, y = Count, color = Cleaned_Keywords, group = Cleaned_Keywords)) +
  geom_line(size = 1.2) +
  geom_point(size = 2) +
  scale_color_manual(values = color_palette) +  # Apply better colors
  labs(title = "Mentions of Key Topics Over Time",
       x = "Year",
       y = "Count of Mentions",
       color = "Topic Keywords") +
  theme_minimal() +
  theme(legend.position = "right", legend.text = element_text(size = 12))  # Improve legend readability

```
@fig-keytopics tracks the mentions of key policy topics—economy, education, energy, health, policy, and tax—in Australian parliamentary speeches from 1998 to 2022. Each topic fluctuates over time, reflecting shifts in legislative focus due to major events.

Economic discussions remain stable but rise post-2008, aligning with the global financial crisis. Tax mentions spike around 2000, likely due to the introduction of the GST, and reappear during key economic reforms. Health surges in 2020, reflecting COVID-19 discussions, while education remains consistently relevant, peaking during policy changes in higher education and funding. Energy mentions rise after 2010, coinciding with climate policy debates like the carbon tax (2012-2014). The steady increase in policy discussions suggests a growing focus on governance and regulatory matters. These trends highlight how parliamentary attention shifts in response to national and global developments such as economic crises, education reforms, climate policies, and the pandemic.
